Discrete Quantum Theories and Computing

### 2018-12-03 dissertation.tex
dissertation.lyx Preface updating...
dissertation.lyx Split Chapter 3 into two chapters...

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newlength{\lyxlabelwidth}      % auxiliary length

@@

\begin{preface*}
\begin{itemize}
\item The latest version of this thesis is in GitHub:
\begin{verbatim}
https://git.io/fxbuG
\end{verbatim}
\item Its source code is typeset using \LyX ~\cite{LyX} in:
\begin{verbatim}
https://git.io/fxbuE
\end{verbatim}
\item Any comments can be left in the Issue part of the GitHub Repository:
\begin{verbatim}
https://git.io/fxbug
\end{verbatim}
\end{itemize}
\end{preface*}

@@

We begin by replacing the continuum of complex numbers by discrete
finite fields in quantum theory. The simplest theory, defined over
unrestricted finite fields, is so weak that it only predict whether
the measure result is possible or impossible. Its computational model
is also weak that it cannot express Deutsch's algorithm, but, paradoxically,
is also so powerful that it can be used to solve the $\usat$ problem,
which is as hard as a general $\NP$-complete problem.

@@

\tableofcontents{}\todo{Split Chapter~\ref{chap:DQT=000026DQC}
into ``Quantum Theories and Computing over Unrestricted Finite Fields''
and ``Quantum Theories and Computing over $\ff{p^{2}}$'' two chapters.
}

@@

\todo{Write a story so that physicists will also be interested in
this thesis. For example, quantum simulation problem and how to apply
these to other physicists work (say two more pages?). Check our previous
papers. Explicitly say what question we wan to solve. In particular,
merging quantum mechanics and computer science is too broad for a
thesis. }The marriage of quantum mechanics and computer science first
envisioned and popularized by Feynman has created an awkward, but
opportune, moment. This embarrassing dilemma was concisely summarized
by Prof.~Aaronson that one of the the following three statements
is false \cite{AaronsonArkhipov2011,BFRDARW2013}:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:Textbook-quantum-mechanics}Textbook quantum mechanics
is correct. 
\item \label{enu:no-efficient-classical-factoring}There does not exist
an efficient classical factoring algorithm. 
\item \label{enu:extended-Church-Turing-Thesis}The extended Church-Turing
thesis — that probabilistic Turing machines can efficiently simulate
any physically realizable model of computation — is correct \cite{Bernstein:1993:QCT:167088.167097,Kaye2007}. 
\end{enumerate}
There is overwhelming evidence to support each of these statements.
The theoretical framework of quantum mechanics \ref{enu:Textbook-quantum-mechanics}
has withstood decades of experimental confirmation. Entire industries
are founded on the assumption \ref{enu:no-efficient-classical-factoring}
that algorithms like RSA are secure and they also have withstood years
of attempted attacks \cite{boneh1999twenty,wiki:RSAFactoring}. Finally
the entire field of complexity theory in computer science which has
also withstood years of field testing rests, in essence, on assumption
\ref{enu:extended-Church-Turing-Thesis} \cite{Aaronson2005,Piccinini2015}.
And yet at least one of these three statements \emph{must be false}!
Indeed if there is a corresponding efficient classical factoring algorithm,
then we concede \ref{enu:no-efficient-classical-factoring}. If it
is correct that Shor's efficient factoring algorithm is realizable,
and we can prove there is no efficient classical factoring algorithm,
then we concede \ref{enu:extended-Church-Turing-Thesis}. Otherwise,
if we cannot implement Shor's algorithm no matter how hard we try,
textbook quantum mechanics, i.e., \ref{enu:Textbook-quantum-mechanics}
may need to be improved by a better theory. It is unlikely that there
will be a simple resolution to this awkward situation. It is more
likely that the resolution will emerge from deep and careful analyses
of the foundations of each field.

@@

\begin{table}
\noindent \centering{}\caption{\label{tab:Comparison}Comparison between the classical computational
models and their physical realizations \todo{Update this table to
include }}
\begin{tabularx}{\linewidth}{ccX}
\toprule 
\addlinespace
 & Model & Physical Realization\tabularnewline\addlinespace
\midrule
\addlinespace
\addlinespace
Continuum & BCSS machine & The precision of an analog computer was limited so that an analog computer may not reproduce the theoretical prediction reliably.\tabularnewline\addlinespace
\addlinespace
\addlinespace
Discrete & Turing machine & A digital computer can be reliably repeated with exact equivalence.\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabularx}
\end{table}
The situations for classical computation is summarized in Table~\ref{tab:Comparison},
and similar situations could potentially be applied to quantum computation.
\todo{Introduction cleaning up until here. }Similar to classical
analog computers manipulating, say, the position of substance, the
most widely used model of quantum computing, the quantum circuit model,
manipulates the probability amplitudes to do the computation. \todo{Play
with some cloud quantum computers to see if I can find more support
evidence. }Because we are agnostic about whether probability amplitude
is quantized or can really be valued in the continuum of complex numbers,
we tried to develop two different types of models: the quantum theories
and computing over finite fields are discrete models, and the quantum
interval-valued probability measures incorporate the idea of measurement
precision into the quantum theory. To distinguish the modified quantum
theories from the usual ones, the usual ones will be called the ``conventional
quantum theory'' where necessary and it will be reviewed in Chapter~\ref{chap:Conventional-Quantum-Theory}~\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}.

Since conventional probability amplitude is in the field of complex
numbers, and the field structure is essential to compute the probability
from some probability amplitudes, we in Chapter~\ref{chap:DQT=000026DQC}
tried to build discrete quantum theories and computing by considering
them over finite fields.

Since error analysis technique cannot always compensate the inevitable
problem of measurement precision, we then incorporate the idea of
finite precision measurement into the quantum theory in Chapter~\ref{chap:QIVPM},
and hope it could describe the physical reality and computability
more faithful.

@@

\chapter{Quantum Theories and Computing over Finite Fields\label{chap:DQT=000026DQC}}

@@

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newlength{\lyxlabelwidth}      % auxiliary length 
\providecommand*{\code}[1]{\texttt{#1}}

@@

\begin{preface*}
\begin{itemize}
\item The latest version of this thesis is in GitHub: \code{https://git.io/fxbuG}
\item Its source code is typeset using \LyX ~\cite{LyX} in: \code{https://git.io/fxbuE}
\item Any comments can be left in the Issue part of the GitHub Repository:
\code{https://git.io/fxbug}
\end{itemize}
\end{preface*}

@@

We begin by replacing the continuum of complex numbers by discrete
finite fields in quantum theory. The simplest theory, defined over
unrestricted finite fields, is so weak that it cannot express Deutsch's
algorithm, but, paradoxically, is also so powerful that it can be
used to solve the $\usat$ problem, which is as hard as a general
$\NP$-complete problem.

Our second framework employs only finite fields over prime numbers
of the form $4\ell+3$, which possess no solutions to $x^{2}+1=0$,
and thus permit an elegant complex representation of the finite field
by adjoining $\rmi=\sqrt{-1}$. Because the states of a discrete $n$-qubit
system are in principle enumerable, we can count the number of states,
and determine the proportions of entangled and unentangled states.
Depending on how we model the measurement process, this improved framework
can be used to implement deterministic Deutsch's algorithm and the
probabilistic Grover search algorithm in a local region, but we still
haven't found a consistent way to treat quantum probability measures
in general.

@@

\tableofcontents{}

@@

\todo{Write a story so that physicists will also be interested in
this thesis. For example, quantum simulation problem and how to apply
these to other physicists work (say two more pages?). Check our previous
papers. Explicitly say what question we want to solve. In particular,
merging quantum mechanics and computer science is too broad for a
thesis. }

The marriage of quantum mechanics and computer science first envisioned
and popularized by Feynman has created an awkward, but opportune,
moment. This embarrassing dilemma was concisely summarized by Prof.~Aaronson
that one of the the following three statements is false \cite{AaronsonArkhipov2011,BFRDARW2013}:
\begin{enumerate}[label=(\roman{enumi})]
\item \label{enu:Textbook-quantum-mechanics}Textbook quantum mechanics
is correct. 
\item \label{enu:no-efficient-classical-factoring}There does not exist
an efficient classical factoring algorithm. 
\item \label{enu:extended-Church-Turing-Thesis}The extended Church-Turing
thesis — that probabilistic Turing machines can efficiently simulate
any physically realizable model of computation — is correct \cite{Bernstein:1993:QCT:167088.167097,Kaye2007}. 
\end{enumerate}
There is overwhelming evidence to support each of these statements.
The theoretical framework of quantum mechanics \ref{enu:Textbook-quantum-mechanics}
has withstood decades of experimental confirmation. Entire industries
are founded on the assumption \ref{enu:no-efficient-classical-factoring}
that algorithms like RSA are secure and they also have withstood years
of attempted attacks \cite{boneh1999twenty,wiki:RSAFactoring}. Finally
the entire field of complexity theory in computer science which has
also withstood years of field testing rests, in essence, on assumption
\ref{enu:extended-Church-Turing-Thesis} \cite{Aaronson2005,Piccinini2015}.
And yet at least one of these three statements \emph{must be false}!
Indeed if there is a corresponding efficient classical factoring algorithm,
then we concede \ref{enu:no-efficient-classical-factoring}. If it
is correct that Shor's efficient factoring algorithm is realizable,
and we can prove there is no efficient classical factoring algorithm,
then we concede \ref{enu:extended-Church-Turing-Thesis}. Otherwise,
if we cannot implement Shor's algorithm no matter how hard we try,
textbook quantum mechanics, i.e., \ref{enu:Textbook-quantum-mechanics}
may need to be improved by a better theory. It is unlikely that there
will be a simple resolution to this awkward situation. It is more
likely that the resolution will emerge from deep and careful analyses
of the foundations of each field.

@@

\begin{table}
\noindent \centering{}\caption{\label{tab:Comparison}Comparison between the classical computational
models and their physical realizations\todo{Put the third row in
the middle alignment.}}
\begin{tabularx}{\linewidth}{ccX}
\toprule 
\addlinespace
 & Discrete & Continuum\tabularnewline\addlinespace
\midrule
\addlinespace
\midrule
\addlinespace
Theoretical  Model & Turing machine & BCSS machine\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
Physical Realization & Digital Computer & Analog Computer\tabularnewline\addlinespace
\addlinespace
\midrule
\addlinespace
How the models realize? & Reliably & Not Reliably:
\begin{itemize}
\item The quality might be quantized.
\item The precision of an analog computer is low.
\end{itemize}\tabularnewline\addlinespace
\bottomrule
\addlinespace
\end{tabularx}
\end{table}
The situations for classical computation is summarized in Table~\ref{tab:Comparison},
and similar situations could potentially be applied to quantum computation.
\todo{Introduction cleaning up until here. }Similar to classical
analog computers manipulating, say, the position of substance, the
most widely used model of quantum computing, the quantum circuit model,
manipulates the probability amplitudes to do the computation. \todo{Play
with some cloud quantum computers to see if I can find more support
evidence. }Because we are agnostic about whether probability amplitude
is quantized or can really be valued in the continuum of complex numbers,
we tried to develop two different types of models: the quantum theories
and computing over finite fields are discrete models, and the quantum
interval-valued probability measures incorporate the idea of measurement
precision into the quantum theory. To distinguish the modified quantum
theories from the usual ones, the usual ones will be called the ``conventional
quantum theory'' where necessary and it will be reviewed in Chapter~\ref{chap:Conventional-Quantum-Theory}~\footnote{Alternative terminology in the literature includes ``actual,'' ``standard,''
and ``ordinary'' quantum theory.}.

\todo{Beginning of Merge Left}Since conventional probability amplitude
is in the field of complex numbers, and the field structure is essential
to compute the probability from some probability amplitudes, we in
Chapter~\ref{chap:DQT=000026DQC} tried to build discrete quantum
theories and computing by considering them over finite fields.

Since error analysis technique cannot always compensate the inevitable
problem of measurement precision, we then incorporate the idea of
finite precision measurement into the quantum theory in Chapter~\ref{chap:QIVPM},
and hope it could describe the physical reality and computability
more faithful.\todo{End of Merge Left}

\todo{Beginning of Merge Right}We begin by replacing the continuum
of complex numbers by discrete finite fields in quantum theory. The
simplest theory, defined over unrestricted finite fields, is so weak
that it only predict whether the measure result is possible or impossible.
Its computational model is also weak that it cannot express Deutsch's
algorithm, but, paradoxically, is also so powerful that it can be
used to solve the $\usat$ problem, which is as hard as a general
$\NP$-complete problem.

Finally, we shift our attention to consider quantum interval-valued
probability measures (IVPMs), which potentially embody both finite
precision measurement and a sensible correspondence to standard quantum
probability. This interval-valued framework not only provides a natural
generalization of both classical IVPMs and conventional quantum probability
measures, but also allows us to establish bounds on the validity of
the Kochen-Specker and Gleason theorems in realistic experimental
environments.\todo{End of Merge Right}

@@

\chapter{Quantum Theories and Computing over Unrestricted Finite Fields}

@@

\chapter{Quantum Theories and Computing over Complexified Finite Fields\label{chap:DQT=000026DQC}}
